\LARGE \bf Approximately Orchestrated Road Traffic Automata:\\ Large-scale Traffic Simulation for Autonomous Vehicles

### 2012-03-14 papers/itsc_2012/final.tex
intermediate version of edited draft

% TODO better name!
\title{\LARGE \bf
Approximately Orchestrated Road Traffic Automata\\
a General Traffic Simulation Framework
}

@@

\begin{abstract} 
Autonomous vehicles have seen great advancements in recent years, and
such vehicles are now ever closer to being commercially available.
The advent of driver-less cars provides opportunities for optimizing
traffic in ways not possible before. This paper introduces an open
source multi-agent traffic simulator called AORTA, which stands for
\textit{Approximately Orchestrated Road Traffic Automata}, and
demonstrates its usage for optimizing autonomous traffic at a
city-wide scale. AORTA creates scale simulations by generating maps
using publicly available road data from \textit{OpenStreetMap}
(OSM). Simulations of specific regions throughout the world can be set
up in the few minutes it takes to export data for the region of
interest from OSM. AORTA focuses on interactions between agents by
defining \textit{behaviors} for individual vehicles and
intersections. These behaviors allow agents to interact with each
other as well as local intersections in any manner required for a
simulation of interest.  This is demonstrated in a simple experiment
for autonomous vehicles presented at the end of this paper.
\end{abstract}

@@

Autonomous vehicle technology has made tremendous progress in the last decade.
In 2004, the farthest distance traveled by a vehicle autonomously in the DARPA
Grand Challenge was $11.9km$ \cite{cnnGrandChallenge2004}. By 2007, six of the
competing teams completed the $96km$ course set for the DARPA Urban Challenge
\cite{spectrumUrbanChallenge2007}. They did so while following the same traffic
laws followed by human drivers, navigating along with other moving vehicles,
and following correct intersection precedence order. Since then, Google's
driver-less cars have clocked more than $250,000km$ on public roads in urban
California, USA \cite{tedThrun2011}. In 2010, researchers from the University
of Parma successfully completed an autonomous \textit{intercontinental} run
from Parma, Italy to Shanghai, China \cite{cnnVislab2010}. The successful
completion of all these milestones suggests that autonomous cars are here to
stay, and are ever closer to becoming commercially available.

The arrival of driver-less cars presents several challenges that need to be
addressed in the near term. Road infrastructure changes could provide important
improvements to traffic involving autonomous vehicles. A number of legal and
safety concerns need to be addressed in the context of driver-less cars. Human
drivers need new methodologies to interact with autonomous drivers on public roads.
This work aims to solve some of these problems by providing a sandbox in which such
methodologies can be found and tested for safety and efficiency before being
deployed in the real world.  In this manner, the present work aids in the effort to find
intermediate stages between entirely human and entirely autonomous traffic.

\begin{figure}[h]
  \centering \includegraphics[width=\linewidth]{downtown_atx.png}
  \caption{A simulation of downtown Austin, Texas.}
  \label{fig:ui_screenshot}
  \vspace{-10pt}
\end{figure}

@@

This paper introduces an open source multi-agent microscopic traffic
simulator called AORTA, which stands for \textit{Approximately
Orchestrated Road Traffic Automata}. The AORTA simulator focuses on
the interactions between individual vehicles and intersections, both
of which act as \textit{agents} in AORTA's framework. Vehicles
interact with other vehicles to pass and change lanes, and interact
with intersections to decide when to cross them. Researchers
developing applications on top of the AORTA simulator define
\textit{behaviors} that these agents follow to interact with one
another. An advantage of such an approach is that by assigning
human-like behaviors such as the ``car following model''
\cite{brackstone1999car} to vehicles, and traffic-signal-like
policies to intersections, AORTA can just as easily simulate human
traffic.

Like any other microscopic traffic simulator, AORTA needs maps to run
simulations. One of AORTA's key features is that it generates maps using road
data available from OpenStreetMap (OSM) \cite{osm}, a Wikipedia-like interface
for world maps. Researchers using AORTA can export any relevant map area from
OSM. This map is then parsed by AORTA to create scale simulations of the real
world. A researcher employing AORTA can have a simulation of any part of the
world for which OSM has data running in about five minutes (allowing for
download time).  AORTA also has the advantage of being open source and easy to
install so that researchers can make any tweaks they find necessary.
\footnote{http://code.google.com/p/road-rage}

After exploring the state of the art in section \ref{sec:related_work} and explaining
the architecture of AORTA (section \ref{sec:arch}) as well as its use of OSM (section \ref{sec:map}),
we will describe simulation in AORTA and how to extend it in sections \ref{sec:simulation} and
\ref{sec:config} respectively. We conclude with an example experiment in section \ref{sec:results}.
In this final experiment, AORTA shows the benefits of an AIM-inspired
\cite{JAIR08-dresner} reservation intersection policy over traditional
traffic signals and stop signs.

@@

Computational processing power has made excellent advancements in the
last two decades. Parallel computing has proved significant for
traffic simulation through the use of Geographic Information System
(GIS) software \cite{pursula1999simulation}. Advanced computing
techniques have enabled microscopic models of traffic simulation to
generate results at a meaningful scale (city-wide or greater). As a
result, a number of good micro-simulators have been introduced in the
past decade. We review a number of such simulators in this section
along with other relevant work.

A number of traffic simulators already make use of OSM data. MATSim is one such
multiagent simulator that focuses on performing large-scale simulations in a
relatively small amount of time \cite{balmer2009matsim}. Traffic demand is
supplied to MATSim in the form of \textit{plans} that an individual may intend
to follow in a given day. MATSim aims to iteratively improve these plans though
offline computation to maximize traffic throughput. In other words, it helps
individuals understand the best way to go about their daily routine to minimize
time spent in transit. This goal is somewhat different from that of AORTA,
where the demand supplied by individuals is taken as is.  Rather than simply improving
routes, AORTA allows entirely different behaviors for both vehicles and intersections.

Another popular open source traffic simulator that can use OSM data is SUMO
\cite{SUMO2011}, which shares the same goals as AORTA and is fairly similar at
a high level. SUMO has been used to study the effect of automated
transportation systems, route planning of individual vehicles, as well as
dynamically adapting traffic signal policies to increase traffic efficiency.
AORTA and SUMO differ in the manner in which intersections are handled. SUMO
uses traditional traffic signals, while AORTA adopts a much more general
\textit{reservation} system. A reservation tells a car when it is safe to take
its desired path through the intersection. Reservations can be allotted to
vehicles through more traditional means such as traffic signals or stop-sign
based intersection precedence, or through more efficient methods designed
specifically for autonomous vehicles \cite{JAIR08-dresner}. Based on experience
with SUMO, AORTA also displays more robust handling of data exported directly from OSM.

@@

A few simulators focus on autonomous vehicles with sensors and actuators
\cite{figueiredo2009approach}. These simulators are used to ensure
that autonomous vehicles are behaving correctly while interacting with
other vehicles and infrastructure. On the other hand, AORTA assumes
that autonomous vehicles have reasonable sensory input and mechanical
output, and instead focuses on interactions between an agent and its
surrounding agents and intersections.

@@

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine \cite{scala}. Scala provides the advantage of functional programming
constructs while still permitting imperative style.  Extensions and clients
built on top of AORTA can be written in either Java or Scala. The software is
open-source \footnote{source code at http://code.google.com/p/road-rage} and
easily extensible. For instance, the stop sign sign policy used in experiments
described in section \ref{sec:results} is implemented in just 70 lines of code.

@@

A map is represented as a directed graph with \emph{lanes} as edges and
\emph{turns} as vertices. A \emph{road} groups one or two sets of lanes
together, one set for each direction of travel.
\emph{Intersections} map the possible turns from incoming to outgoing lanes.
Since both lanes and turns support traffic, they are grouped together as
\emph{traversables}, meaning agents may exist on them.

@@

The builder works by incrementally transforming OSM graphs into AORTA graphs,
with each pass feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so the builder first filters
        these out. Next the builder marks GPS nodes common to multiple roads, since
        they implicitly indicate intersections. Overpasses do not share nodes
        with the roads they cross \cite{osmOverpass}.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item The builder multiplies undirected roads into directed lanes in each
        direction (unless the road is marked one-way). The builder guesses the
        number of lanes based on OSM's ``road type'' tag or an explicit number of
        lanes, when that data is available.
  \item The builder constructs turns between incoming and outgoing lanes at each
        intersection. The angle between the two lanes determines whether the turn is
        a left, right, straight, or U-turn. When several lanes all cross into fewer
        lanes, the builder forces merging of the rightmost lanes. Although
        inaccurate, these patterns are the best compromise with the available data.
  \item Because turns are heuristically placed, the graph is often disconnected.
        In order to ensure that an agent will be able to reach any point in the map,
        the builder removes all but the largest strongly connected component of the
        digraph.  Disconnected lanes are spliced out of larger roads, leaving
        unappealing geometric gaps, but correctly ensuring connectivity for
        corner-case free pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}

>>>>

% TODO better name!
\title{\LARGE \bf
Approximately Orchestrated Road Traffic Automata:\\
Large-scale Traffic Simulation for Autonomous Vehicles
}

@@

\begin{abstract}

Autonomous vehicles have seen great advancements in recent years, and such
vehicles are now ever closer to being commercially available.  The advent of
driverless cars provides opportunities for optimizing traffic in ways not
possible before. This paper introduces an open source multi-agent microscopic
traffic simulator called AORTA, which stands for \textit{Approximately
Orchestrated Road Traffic Automata}, designed for optimizing autonomous traffic
at a city-wide scale. AORTA creates scale simulations of the real world by
generating maps using publicly available road data from \textit{OpenStreetMap}
(OSM). This allows simulations to be set up through AORTA for a desired region
anywhere in the world in a matter of minutes. AORTA allows for optimizing
traffic by creating more intelligent behaviors for individual drivers (agents),
as well as by defining new intersection policies that can be followed by these
autonomous agents. These agent behaviors define how agents interact with one
another, follow intersection policies, and select the next action to take. This
paper demonstrates a simple applications using AORTA through an experiment
testing an intersection policy designed for autonomous vehicles at a city-wide
scale.

\end{abstract}

@@

Autonomous vehicle technology has made tremendous progress in the last decade.
In 2004, the farthest distance traveled by a vehicle autonomously in the DARPA
Grand Challenge was 11.9km \cite{cnnGrandChallenge2004}. By 2007, six of the
competing teams completed the 96km course set for the DARPA Urban Challenge
\cite{spectrumUrbanChallenge2007}. They did so while following the same traffic
laws followed by human drivers, navigating along with other moving vehicles,
and following correct intersection precedence order. Since then, Google's
driverless cars have clocked more than 250,000km on public roads in urban
California, USA \cite{tedThrun2011}. In 2010, researchers from the University
of Parma successfully completed an autonomous \textit{intercontinental} run
from Parma, Italy to Shanghai, China \cite{cnnVislab2010}. The successful
completion of all these milestones suggests that autonomous cars are here to
stay, and are ever closer to becoming commercially available. With the arrival
of autonomous cars, it also becomes possible to optimize traffic in ways not
possible for human drivers.

% The arrival of driverless cars presents several challenges that need to be
% addressed in the near term. Road infrastructure changes could provide important
% improvements to traffic involving autonomous vehicles. A number of legal and
% safety concerns need to be addressed in the context of driverless cars. Human
% drivers need new methodologies to interact with autonomous drivers on public roads.
% This work aims to solve some of these problems by providing a sandbox in which such
% methodologies can be found and tested for safety and efficiency before being
% deployed in the real world.  In this manner, the present work aids in the effort to find
% intermediate stages between entirely human and entirely autonomous traffic.

\begin{figure}[t]
  \centering \includegraphics[width=\linewidth]{downtown_atx.png}
  \caption{A simulation of downtown Austin, Texas.}
  \label{fig:ui_screenshot}
  \vspace{-15pt}
\end{figure}

@@

This paper introduces an open source multi-agent microscopic traffic simulator
called AORTA, which stands for \textit{Approximately Orchestrated Road Traffic
Automata}, a platform designed for testing autonomous vehicle behaviors and
intersection policies. Autonomous vehicles, termed \textit{agents}, use
behaviors to interact with one another, follow intersection policies, and
decide on both long term and short term actions. AORTA allows for the creation
of intersection policies on top of a reservation based intersection manager
\cite{JAIR08-dresner}, and these policies designate when agents can cross the
intersection. AORTA's goal is to allow for the definition of new agent
behaviors and intersection policies to optimize autonomous traffic.
Additionally, by assigning human-like behaviors such as the ``car following
model'' \cite{brackstone1999car} to agents, and traffic-signal-like policies to
intersections, AORTA could potentially simulate human traffic.

Like any other microscopic traffic simulator, AORTA needs maps to run
simulations. One of AORTA's key features is that it generates maps using real
road data available from OpenStreetMap (OSM) \cite{osm}, a Wikipedia-like
interface for world maps. A map for any desired city in the world can be
downloaded, which is then parsed by AORTA to set up a scale simulation of the
real world in a few minutes. AORTA is available open-source and is easily
extensible \footnote{Code available at
\texttt{http://code.google.com/p/road-rage}} \commentpk{could also change this
to a citation}, making it easier for users to test out a number of agent
behaviors and intersection policies.

This paper explores the state-of-the-art in traffic simulators in section
\ref{sec:related_work}, followed by a description of AORTA's architecture in
section \ref{sec:arch}.  Use of OSM data in AORTA is explained in section
\ref{sec:map}, along with a description of the simulator and the default
behaviors available in AORTA in section \ref{sec:simulation}. An experiment
investigating the benefit of an alternate intersection policy designed for
autonomous vehicles is presented in \ref{sec:results}. The paper then concludes
with a discussion on future work.

@@

Computational processing power has made excellent advancements in the last two
decades. Parallel computing has proved significant for traffic simulation
through the use of Geographic Information System (GIS) software
\cite{pursula1999simulation}. \commentpk{preceding sentence doesnt paraphrase
the ref correctly. double check}. Advanced computing techniques have enabled
microscopic models of traffic simulation to generate results at a meaningful
scale (city-wide or greater). As a result, a number of good micro-simulators
have been introduced in the past decade. We review a number of such simulators
in this section along with other relevant work.

A number of traffic simulators already make use of OSM data. MATSim is one such
multiagent simulator that focuses on performing large-scale simulations in a
relatively small amount of time \cite{balmer2009matsim}. Traffic demand is
supplied to MATSim in the form of \textit{plans} that an individual may intend
to follow in a given day. MATSim aims to iteratively improve these plans though
offline computation to maximize traffic throughput. In other words, it helps
individuals understand the best way to go about their daily routine to minimize
time spent in transit. This goal is somewhat different from that of AORTA,
where the demand supplied by individuals is taken as is. Rather than simply
improving routes, AORTA allows creating entirely different behaviors for both
vehicles and intersections.

Another popular open source traffic simulator that can use OSM data is SUMO
\cite{SUMO2011}, which shares the same goals as AORTA and is fairly similar at
a high level. SUMO has been used to study the effect of automated
transportation systems, route planning of individual vehicles, as well as
dynamically adapting traffic signal policies to increase traffic efficiency.
AORTA and SUMO differ in the manner in which intersections are handled. SUMO
uses traditional traffic signals, while AORTA adopts a much more general
\textit{reservation} system. A reservation tells a car when it is safe to take
its desired path through the intersection. Reservations can be allotted to
vehicles through more traditional means such as traffic signals or stop-sign
based intersection precedence, or through more efficient methods designed
specifically for autonomous vehicles \cite{JAIR08-dresner}. Based on experience
with SUMO, AORTA also displays more robust handling of data exported directly
from OSM. \commentpk{I am worried about the last line. It might be true, but it
is a fairly strong claim - we should put a citation of us handling something
that has been documented as not being handled via SUMO}

@@

Simulators have also focused on autonomous vehicles with sensors and actuators
\cite{figueiredo2009approach}. These simulators are used to ensure that
autonomous vehicles are behaving correctly while interacting with other
vehicles and infrastructure, given their configuration of sensors and
actuators. On the other hand, AORTA assumes that autonomous vehicles have
reasonable sensory input and mechanical output, and instead focuses on
interactions between an agent and its surrounding agents and intersections.

@@

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine \cite{scala}. Scala provides the advantage of functional programming
constructs while still permitting imperative style.  Extensions and clients
built on top of AORTA can be written in either Java or Scala. The software is
open-source and easily extensible. For instance, the stop sign sign policy used
in experiments described in section \ref{sec:results} is implemented in just 70
lines of code.

Section \ref{sec:map} and \ref{sec:simulation} describe map construction from
OSM data and the simulation engine used in the AORTA framework respectively.

@@

A map is represented as a directed graph with \emph{lanes} as edges and
\emph{turns} as vertices. A \emph{road} groups a set of lanes together, one set
for each direction of travel. \emph{Intersections} map the possible turns from
incoming to outgoing lanes.  Since both lanes and turns support traffic, they
are grouped together as \emph{traversables}, meaning agents may exist on them.

@@

The builder works by incrementally transforming OSM graphs into AORTA graphs,
with each pass feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so the builder first filters
        these out. Next, the builder marks GPS nodes common to multiple roads, since
        they implicitly indicate intersections. Overpasses do not share nodes
        with the roads they cross \cite{osmOverpass}.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item The builder multiplies undirected roads into directed lanes in each
        direction (unless the road is marked one-way). The builder guesses the
        number of lanes based on OSM's ``road type'' tag or an explicit number of
        lanes, when that data is available.
  \item The builder constructs turns between incoming and outgoing lanes at each
        intersection. The angle between the two lanes determines whether the turn is
        a left, right, straight, or U-turn. When several lanes all cross into fewer
        lanes, the builder forces merging of the rightmost lanes. Although
        inaccurate, these patterns are the best compromise with the available data.
  \item Because turns are heuristically placed, the graph is often disconnected.
        In order to ensure that an agent will be able to reach any point in the map,
        the builder removes all but the largest strongly connected component of the
        digraph.  Disconnected lanes are spliced out of larger roads, leaving
        unappealing geometric gaps, but correctly ensuring connectivity for
        corner-case free pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}
